---
author: Balint Mazzag
categories:
- Maps
- Shiny
- Travel
date: "2022-12-23"
draft: false
excerpt: A Shiny application which maps out the countries where I have travelled so far, and how I rank them.
featured: true
layout:
links:
- icon: door-open
  icon_pack: fas
  name: website
  url: https://balint-mazzag.shinyapps.io/wherehaveibeen/
- icon: github
  icon_pack: fab
  name: code
  url: https://github.com/mbalint9901/wherehaveibeen/
subtitle: ""
tags:
- hugo-site
title: Where Have I Been?
output: html_document
---

### A Shiny application which maps out the countries where I have travelled so far, and how I rank them.

---

Since I was very young, I kept track of everything, related to geography, or countries in general. I always enjoyed looking at maps and visualising travel between them. In the past 23 years, I had the opportunity to travel to many of them, and I rank them based on my personal preference.

---

The embedded Shiny application deployed to visualize the tiers of countries
*As the datatable is not dynamically sizeable, it may appear way larger on your device then it should on the original app*

```{r, echo=FALSE}
knitr::include_app("https://balint-mazzag.shinyapps.io/wherehaveibeen/", height = "700px")
```


> ##### Disclaimer
>
> The tier list was assembled based on preference of the country's weather, cleanliness, cost of living and other important factors that was considered during either vacation or living in a certain location.



### Build-up of the application

#### Step 1 - Creating the data for ranking and visualising

```{r setup, include=FALSE}

library(tidyverse)
library(leaflet)
library(readxl)

```

Although, I usually read in Excel files with the `readxl` package, in the blog, I will initialise my raw 
dataframe from the GitHub repo of the project.

- `data` is the initial raw dataframe
- `data_geom` includes the shapefile of the countries
- `factpal` is the coloring palette that I will use for the different Tiers


```{r data_input}

#readxl::read_xlsx("/data/country_year_tier.xlsx") #Works for local projects

load(url("https://github.com/mbalint9901/wherehaveibeen/blob/main/data/data_render.RData?raw=true"))
ls()

```

In these dataframes, I already set up the Tier list and also the shapefiles for plotting them later on.
As it is a Shiny application, there are two files that go hand-in-hand separately:

- `ui.R` which handles the user interface and the reactive operations
- `server.R` that handles recalculation and rendering of the plot

Here, I can only show you the static version of the map, however it is the basic logic behind the application

For maps in Shiny I usually opt for the nicer `leaflet` package, which renders layers on top of each other:
1. First, the `addProviderTiles` renders the background map
2. Then, the `addPolygons` feature adds the shapefile layers from the dataframe to the leaflet map
3. We can tailor the formatting of the polygons to our needs and desires, with `highlightOptions` and
the popup labels and markers

I just tried out how to insert flag images into the popup labels, through URLs pointing to an online resource 
of images of flags. Although it takes a little more time to load the images, it is easily implmeneted in any other project
without having to download a whole package of *SVG* images

```{r data_analyse, warning=FALSE, message=FALSE, include = FALSE}
library(leaflet)
library(sf)
library(shiny)
```


To make it interactive, and to mimic the behaviour of the application, we can create a filter for the 
year I travelled to a destination.


```{r data_render, echo = FALSE, warning=FALSE, message=FALSE, results='hold'}


first_visit_var <- 2022

data_geom %>%
  st_as_sf() %>% 
  filter(first_visit <= first_visit_var) %>%
  leaflet() %>%
  setView(lng = -50, lat = 30, zoom = 2) %>%
  # addTiles() %>%
  addProviderTiles(providers$CartoDB.PositronNoLabels) %>%
  addPolygons(color = ~factpal(tier), weight = 1,
              layerId = data$country_code,
              smoothFactor = 0.5, fillOpacity = 0.8,
              opacity = 1,
              highlightOptions = highlightOptions(color = "white", weight = 2,
                                                  bringToFront = TRUE),
              popup= ~paste0(
                '<strong>', country ,
                '</strong><br/>Tier:', tier,
                '<br/>First visited:', first_visit,
                '<br/><a href = "https://www.wikipedia.org/wiki/',
                country,
                '"> Wikipedia link </a>'),
              label =
                ~paste0(
                  "<img src='",image_link,"' height='24px'>",
                  "&nbsp;&nbsp;&nbsp;&nbsp;<strong>", country, "</strong><br/>Click for more info!"
                ) %>% lapply(htmltools::HTML))


```

```{r data_render, eval=FALSE, echo=T}
```
And voilá, my own travel diary is starting to take shape, but for interactivity, check my code on
Github

